---
title: "RWorksheet_Salvador#6"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.Create a data frame for the table below. Show your solution.
```{r}

scores <- data.frame(
  Student = c(1:10),
  PreTest = c(55, 54, 47, 57, 51, 61, 57, 54, 63, 58),
  PostTest = c(61, 60, 56, 63, 56, 63, 59, 56, 62, 61)
)  
scores

```
#1a.Compute the descriptive statistics using different packages (Hmisc and pastecs). Write the codes and its result.
```{r}
# Install and load the Hmisc package
install.packages("Hmisc")
library(Hmisc)

# I create a data frame named "scores".
scores <- data.frame(
  Student = c(1:10),
  PreTest = c(55, 54, 47, 57, 51, 61, 57, 54, 63, 58),
  PostTest = c(61, 60, 56, 63, 56, 63, 59, 56, 62, 61)
)

#compute descriptive statistics using hmisc
summary_hmisc <- describe(scores)
summary_hmisc

# install pastecs
install.packages("pastecs")
library(pastecs)

# Compute descriptive statistics using pastecs
summary_pastecs <- stat.desc(scores)
summary_pastecs
```
#2a. The Department of Agriculture was studying the effects of several levels of a fertilizer on the growth of a plant. For some analyses, it might be useful to convert the fertilizer levels to an ordered factor.
```{r}
# this is the sample data given.
data <- c(10, 10, 10, 20, 20, 50, 10, 20, 10, 50, 20, 50, 20, 10)

# this is were i convert fertilizer levels to an ordered factor.
ferti_lvls <- ordered(data, levels = c(10, 20, 50))

# this is were i display the factor
ferti_lvls

#This code shows the different levels of fertilizer used in a plant study and provides a summary of the data. You can change the order of the levels as needed.
```
#3. Abdul Hassan, president of Floor Coverings Unlimited, has asked you to study the exercise levels undertaken by 10 subjects were “l”, “n”, “n”, “i”, “l” , “l”, “n”, “n”, “i”, “l” ; n=none, l=light, i=intense 
#a. What is the best way to represent this in R?

```{r}
# i create a factor variable for exercise levels
exer_lvls <-  factor(exercise_levels, levels = c("n", "l", "i"), labels = c("none", "light", "intense"))


# i display the factor variable
exer_lvls

#The most effective way to show the exercise levels of the 10 subjects is by using a factor variable in R. Each intensity level of exercise ("none," "light," "intense") can be portrayed as a different factor level.
```
#4a. Sample of 30 tax accountants from all the states and territories of Australia and their individual state of origin is specified by a character vector of state mnemonics as:
#state <- c("tas", "sa", "qld", "nsw", "nsw", "nt", "wa", "wa", "qld",
#"vic", "nsw", "vic", "qld", "qld", "sa", "tas", "sa", "nt",
#"wa", "vic", "qld", "nsw", "nsw", "wa", "sa", "act", "nsw",
#"vic", "vic", "act")
#a. Apply the factor function and factor level. Describe the results.

```{r}
# this is the sample data given
state <- c("tas", "sa", "qld", "nsw", "nsw", "nt", "wa", "wa", "qld",
  "vic", "nsw", "vic", "qld", "qld", "sa", "tas", "sa", "nt",
  "wa", "vic", "qld", "nsw", "nsw", "wa", "sa", "act", "nsw",
  "vic", "vic", "act")

# this is a factor variable
state_factor <- factor(state)

# display the factor levels
levels(state_factor)

#The result you see when using levels(state_factor) shows the specific categories or states that are part of the factor variable called state_factor.
```

