---
title: "RWorksheet_Salvador#6"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.Create a data frame for the table below. Show your solution.
```{r}

scores <- data.frame(
  Student = c(1:10),
  PreTest = c(55, 54, 47, 57, 51, 61, 57, 54, 63, 58),
  PostTest = c(61, 60, 56, 63, 56, 63, 59, 56, 62, 61)
)  
scores

```
#1a.Compute the descriptive statistics using different packages (Hmisc and pastecs). Write the codes and its result.
```{r}
# Install and load the Hmisc package
install.packages("Hmisc")
library(Hmisc)

# I create a data frame named "scores".
scores <- data.frame(
  Student = c(1:10),
  PreTest = c(55, 54, 47, 57, 51, 61, 57, 54, 63, 58),
  PostTest = c(61, 60, 56, 63, 56, 63, 59, 56, 62, 61)
)

#compute descriptive statistics using hmisc
summary_hmisc <- describe(scores)
summary_hmisc

# install pastecs
install.packages("pastecs")
library(pastecs)

# Compute descriptive statistics using pastecs
summary_pastecs <- stat.desc(scores)
summary_pastecs
```
#2a. The Department of Agriculture was studying the effects of several levels of a fertilizer on the growth of a plant. For some analyses, it might be useful to convert the fertilizer levels to an ordered factor.
```{r}
# this is the sample data given.
data <- c(10, 10, 10, 20, 20, 50, 10, 20, 10, 50, 20, 50, 20, 10)

# this is were i convert fertilizer levels to an ordered factor.
ferti_lvls <- ordered(data, levels = c(10, 20, 50))

# this is were i display the factor
ferti_lvls

#This code shows the different levels of fertilizer used in a plant study and provides a summary of the data. You can change the order of the levels as needed.
```
#3. Abdul Hassan, president of Floor Coverings Unlimited, has asked you to study the exercise levels undertaken by 10 subjects were “l”, “n”, “n”, “i”, “l” , “l”, “n”, “n”, “i”, “l” ; n=none, l=light, i=intense 
#a. What is the best way to represent this in R?

```{r}
# i create a factor variable for exercise levels
exer_lvls <-  factor(exercise_levels, levels = c("n", "l", "i"), labels = c("none", "light", "intense"))


# i display the factor variable
exer_lvls

#The most effective way to show the exercise levels of the 10 subjects is by using a factor variable in R. Each intensity level of exercise ("none," "light," "intense") can be portrayed as a different factor level.
```
#4a. Sample of 30 tax accountants from all the states and territories of Australia and their individual state of origin is specified by a character vector of state mnemonics as:
#state <- c("tas", "sa", "qld", "nsw", "nsw", "nt", "wa", "wa", "qld",
#"vic", "nsw", "vic", "qld", "qld", "sa", "tas", "sa", "nt",
#"wa", "vic", "qld", "nsw", "nsw", "wa", "sa", "act", "nsw",
#"vic", "vic", "act")
#a. Apply the factor function and factor level. Describe the results.

```{r}
# this is the sample data given
state <- c("tas", "sa", "qld", "nsw", "nsw", "nt", "wa", "wa", "qld",
  "vic", "nsw", "vic", "qld", "qld", "sa", "tas", "sa", "nt",
  "wa", "vic", "qld", "nsw", "nsw", "wa", "sa", "act", "nsw",
  "vic", "vic", "act")

# this is a factor variable
state_factor <- factor(state)

# display the factor levels
levels(state_factor)

#The result you see when using levels(state_factor) shows the specific categories or states that are part of the factor variable called state_factor.
```
#5. From #4 - continuation:
• Suppose we have the incomes of the same tax accountants in another vector (in suitably
large units of money)
incomes <- c(60, 49, 40, 61, 64, 60, 59, 54,
62, 69, 70, 42, 56, 61, 61, 61, 58, 51, 48,
65, 49, 49, 41, 48, 52, 46, 59, 46, 58, 43)
#5a. Calculate the sample mean income for each state we can now use the special function
tapply():
Example: giving a means vector with the components labelled by the levels
incmeans <- tapply(incomes, statef, mean)
Note: The function tapply() is used to apply a function, here mean(), to each group
of components of the first argument, here incomes, defined by the levels of the second
component, here state 2
• 2
that tapply() also works in this case when its second argument is not a factor,
• e.g., ‘tapply(incomes, state)’, and this is true for quite a few other functions, since
arguments are coerced to factors when necessary (using as.factor()).

```{r}
incomes <- c(60, 49, 40, 61, 64, 60, 59, 54,
  62, 69, 70, 42, 56, 61, 61, 61, 58, 51, 48,
  65, 49, 49, 41, 48, 52, 46, 59, 46, 58, 43)


# this line of my code calculate the sample mean income for each state using tapply.
inc_means <- tapply(incomes, state_factor, mean)

# Display the results
inc_means
```
#5b

#this is the output that shows the average income of each state.
#act      nsw       nt      qld       sa      tas      vic       wa 
#44.50000 57.33333 55.50000 53.60000 55.00000 60.50000 56.00000 52.25000

#The tapply function helps us in determining the average income of tax accountants in each state. The output provides us with the average income for tax accountants in different states.

#6. Calculate the standard errors of the state income means (refer again to number 3) stdError <- function(x) sqrt(var(x)/length(x))
#Note: After this assignment, the standard errors are calculated by: incster <- tapply(incomes, statef, stdError)

#6a.What is the standard error? Write the codes.

```{r}
stdError <- function(x) sqrt(var(x) / length(x))

inc_std_errors <- tapply(incomes, state_factor, stdError)

print(inc_std_errors)
```
#6b. Interpret the result.
These numbers show how much the average incomes in each place might change. If the numbers are higher, we're not so sure about the average incomes. The "stdError function" figures out these numbers, helps us know how sure we can be about the average incomes for tax accountants in different places.

#7. Use the titanic dataset.

#a. subset the titatic dataset of those who survived and not survived. Show the codes and its result.

```{r}
install.packages("titanic")
library(titanic)

data("titanic_train")

survived <- subset(titanic_train, Survived == 1)

not_survived <- subset(titanic_train, Survived == 0)

head(survived)
head(not_survived)

#These commands helps examine the data for passengers who either survived or did not survive the Titanic disaster. 
```

